---
title: "Survival Functions: Kaplan-Meier, Fleming-Harrington"
author: "Brendan Mapes"
date: "11/19/2021"
output: html_document
---
_Estimating the survival curves for time by SEX using the following methods:Kaplan-Meier, Fleming-Harrington._

```{r}
library(survival)
library(dplyr)
library(magrittr)
library(survminer)
lung <- data(cancer)
lung <- cancer %>% select(inst, time, status, age, sex)
kmfit <- survfit(Surv(time, status) ~ sex, data=lung)
ggsurvplot(
    fit = kmfit, 
    xlab = "Days", 
    ylab = "Overall survival probability")
```
Important to note on these plots, sex 1 refers to male, and sex 2 refers to female. Line colors here aren't assigned as one may expect. Above is the plot from Kaplan-Meier method, and below is plot from the Fleming-Harrington method.
```{r}
fhfit <- survdiff(Surv(time, status) ~ sex, data=lung, rho=0)
fhfit2 <- survfit(Surv(time, status) ~ sex, data=lung, type="fleming-harrington")
ggsurvplot(
    fit = fhfit2, 
    xlab = "Days", 
    ylab = "Overall survival probability")
```

```{r}
kmfit
```

```{r}
fhfit2
```
_Estimating the median survival time, using the estimated survival curves._
Median survival rates on the plot can be found by locating the point on the x-axis that corresponds to 0.5 probability of survival on the y-axis on the survival curve. More simply in R, we can look to the outputs above for median estimates. The Kaplan-Meier method results in estimates of median survival of 270 days for males and 426 days for females. Fleming-Harrington gives estimates of 218 days for males and 348 days for females.

_Using log-rank test to compare the survival distributions for Male and Female_
```{r}
log_rank <- survdiff(Surv(time, status) ~ sex, data = lung)
log_rank
```
The log rank test with survdiff() suggests that the males live longer on average, although there are more males in the dataset than there are females. 

